\phantomsection
\begin{Acknowledgements}
	I am deeply indebted to many amazing people in life, but I foremost wish to express my highest appreciation to my wonderful wife, \textbf{Shafta}. My dearest wife, your continued love, support, and motivation helped me through the challenging years of my graduate study. You always stood beside me through twists and turns; 	good and overwhelming times; wonderful and hard moments. Most of all, you created a life and made me a father. Ever since, I admire you more for how you are raising our child and fostering confidence, love, and humane values in him. Life's greatest gift is having you as my other half and the mother of my child. 
	
	Next, I am extremely grateful and privileged to have \textbf{Prof. Marino Arroyo} as my advisor.  He introduced me to the wonderful world of Computational Biophysics and with his vision steered my project to explore the awe-inspiring microscopic wonders of biological systems. His contribution to my research has been immense both in terms of 
	substance and time. His superpowers of great technical mentorship, critical feedback, patience, empathy, and compassion helped me to grow as a scientist, as well as, as a person.  The success and completion of my dissertation would not have been possible without the continual support and mentorship of \textbf{Alejandro Torres} and \textbf{Guillermo Vilanova}. Their technical skillset and patience in working with me through the nitty gritty challenges of my research allowed me to foster an interdisciplinary skill set on the intersection of mathematical modeling, computational science, and biology. I can write at length, but no amount of words in my acknowledgment can do any justice to the enormous contribution you two have made to my research and learning experience.
	
	
	I would also like to thank my polymath colleague \textbf{Marco De Corato} who instilled in me a passion for theoretical physics. In particular, I learned to solve complex models with rather simple analytical techniques, a skill that was essential in shaping the first part of my dissertation.  I would further like to acknowledge the support of my brilliant colleague, \textbf{Marco Pensalfini}. He exposed me to the realm of discrete network simulations and our collaboration gave a microscopic context to my broader research.
	Furthermore, I have always enjoyed discussing with both Marcos the technical and non-technical aspects of being a good researcher and consolidating a career in today’s academia. 
	It would be not fair if I omit to acknowledge the role of my laboratory family, i.e,  my cruel \textit{amiga para siempre} \textbf{Chiara Venturini}, hilarious \textbf{Jordi Font},  certified cryptocurrency expert \textbf{Hossein Mohammadi}, witty \textbf{Adam Ouzeri}, brilliant \textbf{David Codony} and lively \textbf{Caterina Tozzi} in making \textit{room 208} a dynamic and fun place for research and learning.  
	
	Lastly, I wish to express gratitude toward the \textbf{La Caixa Foundation} and the \textbf{European Research Council} for generously funding my graduate education, organizing training programs to enhance my technical and leadership qualities, and financing one of the highest pleasures of graduate life, i.e., conference travels which allowed me to expose myself to new ideas, as well as, disseminate my own.
	
	
\end{Acknowledgements}
\cleardoublepage